---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

devtools::load_all('.')
library(testthis)
library(rstudioapi)
```

## Overview

Testthis provides several tools to make unit testing in R slightly more 
comfortable. It is designed to be used in conjunction with `testthat` and 
`devtools` packages. 

## Rstudio addins (requires Rstudio)

Provides Rstudio addins that enable to assign hotkeys to menial tasks. I 
recommend assigning `ctrl+alt+Insert`, `ctrl+alt+Pos1` and `ctrl+alt+PageUp`.

* `test_this`: Tests the currently open file.
* `lest_this`: "Load and test"; first calls `devtools::load_all()`, and then 
  test the currently open file
* `open_tests`: Opens the associated testfile in an editor window. If the 
  currently open file already is a testfile, it opens the associated file in
  the `/R` directory.

The function above assume that if the current filename is `currentfile.R`,
the associated test file is `/tests/testthat/test_currentfile.R`. If you want
to modify this behaviour you can put the tag `#* @testfile anotherfile` anywhere
in your code, usually the top or bottom of your .R file.

The following example will associate the file `/tests/testthat/other_test_file.R`
instead of `/tests/testthat/test_open_tests.R` with `R/open_tests`:
```{r eval = FALSE}
# file R/open_tests.R
#* @testfile other_testfile

open_tests <- function(){
  fname <- get_testfile_name()

  if(file.exists(fname)){
    rstudioapi::navigateToFile(fname)
  } else {
    test_skeleton(fname, open = TRUE)
  }
}
```

## Test coverage checklist

Testthis now provides a checklist like test coverage analyser. This is 
non-automatic and requires you to manually mark functions as tested in your 
code via special comments. If you look for automatic test coverage analysis, 
you might want to check out the `covr` package instead.

For testthis to recognize wheter a unit test exists for a function, the function
name must either be mention in a `testthat()` `desc` argument, or marked via
special comment tag: `#* @testing functionname`. The following example will mark
the functions `parse_testthis_comments` and `detect_testthis_comments` as 
tested:

```{r, eval = FALSE}
# file /tests/testhis/parse_testthis_comments.R

test_that("parse_testthis_comments works as expected", {
  #* @testing detect_testthis_comments
  ...
}
```

The test_coverage summary for a package may look like this (+ marks function for
which tests exist):

```{r}
get_test_coverage(pkg = system.file(package = 'testthis'))
```

## Testthis Tags

Test_this tags are special comments that modify the behavious of the functions
supplied by this package. They are of the Form `#* @tag <value>`. Please not 
that only some test_this tags really require a `<value>`. The tag system should
be considered to be in an alpha-state.

### `pkg/R/*.R`

* `@testfile <filename>`: manually specifiy associated test file. Should 
  usually start with `test_`. This is used by `test_this()`, `lest_this()` and
  `open_tests()`.

### `pkg/tests/testthat/test_*.R`

* `@testing <functionname>`:  mark `functionname` as tested.
  Should usually go next the associated `test_that()` call. This is used by
  `get_test_coverage()`.
* `@skip`: skip test when running `testthis::test_with_skip()`. Best placed 
  directly below the `context()` call.

also look at `/test/testthat/test_Taglist.R` for examples.


Package options
---------------

You can set the following global options using `options()`:

* `testthis.sep`: Default separator to use when creating test files with 
  `test_skeleton()`. Defaults to `_`, must be either `_` or `-`. I.e whether
  you want your files to be named `test_foofunction.R` or `test-foofunction.R`
