---
title: "testthis - package vignette"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{testthis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

```{r}
library(testthis)
```
Testthis provides several tools to make unit testing in R slightly more 
confortable. It is designed to be used in conjunction with `testthat` and 
`devtools` packages. 

## Rstudio addins (requires Rstudio)

Provides Rstudio addins that enable to assign hotkeys to menial tasks. I 
recommend assigning `ctrl+alt+Insert`, `ctrl+alt+Pos1` and `ctrl+alt+PageUp`.

* *test_this*: Tests the currently open file.
* *lest_this*: "Load and test"; first calls `devtools::load_all`, and then test 
                the currently open file
* *open_tests*: Opens the associated testfile in an editor window. If the file 
                does not exists, it is created with a minimal testing 
                infrastructure skeleton.
                
All those functions assume that if the current filename is `currentfile.R`,
the associated test file is `/tests/testthat/test_currentfile.R`. If you want
to modify this behaviour you can put the tag `#* testfile anotherfile` anywhere
in your code, usually the top or bottom of your .R file.

The following example will associate the file `/tests/testthat/other_test_file.R`
instead of `/tests/testthat/test_open_tests.R` with `R/open_tests`:
```{r}
# file R/open_tests.R
#* @testfile other_testfile

open_tests <- function(){
  fname <- get_testfile_name()

  if(file.exists(fname)){
    rstudioapi::navigateToFile(fname)
  } else {
    test_skeleton(fname, open = TRUE)
  }
}
```

## Test coverage checklist

For testthis to recognize wheter a unit test exists for a function, the function
name must either be mention in a `testthat()` `desc` argument, or marked via
special comment tag: `#* @testing functionname`. The following example will mark
the functions `parse_testthis_comments` and `detect_testthis_comments` as 
tested:

```{r, eval = FALSE}
# file /tests/testhis/parse_testthis_comments.R

test_that("parse_testthis_comments works as expected", {
  #* @testing detect_testthis_comments
  ...
}
```

The test_coverage summary for a package may look like this (+ marks function for
which tests exist):

```{r}
get_test_coverage(pkg = '~/Dropbox/workspace/r/testthis')
```
