---
title: "testthis - package vignette"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{testthis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

```{r include = FALSE}
library(testthis)
library(rstudioapi)
```
Testthis provides several tools to make unit testing in R slightly more 
confortable. It is designed to be used in conjunction with `testthat` and 
`devtools` packages. 

## Rstudio addins (requires Rstudio)

Provides Rstudio addins that enable to assign hotkeys to menial tasks. I 
recommend assigning `ctrl+alt+Insert`, `ctrl+alt+Pos1` and `ctrl+alt+PageUp`.

* `test_this`: Tests the currently open file.
* `lest_this`: "Load and test"; first calls `devtools::load_all()`, and then test 
                the currently open file
* `open_tests`: Opens the associated testfile in an editor window. If the file 
                does not exists, it is created with a minimal testing 
                infrastructure skeleton.
                
All those functions assume that if the current filename is `currentfile.R`,
the associated test file is `/tests/testthat/test_currentfile.R`. If you want
to modify this behaviour you can put the tag `#* testfile anotherfile` anywhere
in your code, usually the top or bottom of your .R file.

The following example will associate the file `/tests/testthat/other_test_file.R`
instead of `/tests/testthat/test_open_tests.R` with `R/open_tests`:
```{r eval = FALSE}
# file R/open_tests.R
#* @testfile other_testfile

open_tests <- function(){
  fname <- get_testfile_name()

  if(file.exists(fname)){
    rstudioapi::navigateToFile(fname)
  } else {
    test_skeleton(fname, open = TRUE)
  }
}
```

## Test coverage checklist

For testthis to recognize wheter a unit test exists for a function, the function
name must either be mention in a `testthat()` `desc` argument, or marked via
special comment tag: `#* @testing functionname`. The following example will mark
the functions `parse_testthis_comments` and `detect_testthis_comments` as 
tested:

```{r, eval = FALSE}
# file /tests/testhis/parse_testthis_comments.R

test_that("parse_testthis_comments works as expected", {
  #* @testing detect_testthis_comments
  ...
}
```

The test_coverage summary for a package may look like this (+ marks function for
which tests exist):

```{r}
get_test_coverage(pkg = '~/Dropbox/workspace/r/testthis')
```

## Test This Tags

Test_this tags are special comments that modify the behavious of the functions
supplied by this package. They are of the Form `#* @tag <value>`. Please not 
that only some test_this tags really require a `<value>`. The tag system should
be considered to be in an alpha-state.

### `pkg/R/*.R`

* `@testfile <filename>`: manually specifiy associated test file. Should 
  usually start with `test_`. This is used by `test_this()`, `lest_this()` and
  `open_tests()`.

### `pkg/tests/testhat/test_*.R`

* `@testing <functionname>`:  mark `functionname` as tested.
  Should usually go next the associated `test_that()` call. This is used by
  `get_test_coverage()`.
* `@skip`: skip test when running `test_this::test()`. Best placed directly
  below the `context()` call.

also look at `/test/testthat/test_Taglist.R` for examples.
