---
title: "testthis package vignette"
author: "Stefan Fleck"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{testthis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

devtools::load_all('.')
library(testthis)
library(rstudioapi)
```

## Overview

Testthis contains tools to make unit testing in R more comfortable. It is 
designed to complement the packages **testthat**, **devtools** and **usethis**. 



## RStudio addins 

Testthis provides RStudio addins for common tasks. These can be called like
normal R function, but it is also possible to assigned to hotkeys to them in
RStudio (*Tools/Modify Keyboard Shortcuts*). I recommend assigning 
`ctrl+alt+Insert`, `ctrl+alt+Pos1` and `ctrl+alt+PageUp` to the following 
functions:

* `test_this()`: Run tests associated with the currently open R script file.
* `lest_this()`: "Load and test"; As above, but call `devtools::load_all()` first
* `open_testfile()`: Opens the associated testfile in an editor window. If the 
  currently open file already is a testfile, it opens the associated file in
  the `/R` directory. Can be used to jump back and forth between both.
* `test_with_skip()`: Like `devtools::test()`, but does not run test files that
  contain the line `#' @skip.`

The functions above assume that if the current filename is `currentfile.R`,
the associated test file is `/tests/testthat/test_currentfile.R`. If you want
to modify this behaviour you can put the tag `#* @testfile anotherfile` anywhere
in your code, usually the top or bottom of your `.R` file. Each R script can only
have a single associated testfile.

The following example will associate the file `/tests/testthat/other_test_file.R`
instead of `/tests/testthat/test_open_testfile.R` with `R/open_testfile`:
```{r eval = FALSE}
# file R/open_testfile.R

#* @testfile other_testfile
open_testfile <- function(){
  fname <- get_testfile_name()

  if(file.exists(fname)){
    rstudioapi::navigateToFile(fname)
  } else {
    test_skeleton(fname, open = TRUE)
  }
}
```




## Managing test data

Sometimes, it might become necessary to have access to complex data sets for
tests. The recommended way for dealing with this is to provide an example 
dataset in your packages `data/` directory; however, there are also usecases 
where you want seperate data files for your tests.

Testthis provides three functions to to make dealing with separate test data 
files a bit easier:

* `use_testdata()` places a single R object in the `tests/testhat/testdata` 
  directory. It is analoguous to `usethis::use_data()`, except that it saves the
  Object in the `.rds` format, which is more convenient for single R Objects
  than `.rda` or `.Rdata` (see `?readRDS`). 
* `use_testdata_raw()` creates the directory `tests/testhat/testdata-raw`. Use
  this directory to put scripts that generate the data in 
  `tests/testhat/testdata`. 
* `read_testdata()` is a simple wrapper for `readRDS()` to read files
  in `tests/testhat/testdata`. 


```{r, eval = FALSE}
use_testdata(iris)
## Saving to 
## * r/testthis/tests/testthat/testdata/iris.rds

tdat <- read_testdata("iris.rds")

use_testdata_raw()
## ✔ Creating 'tests/testthat/testdata-raw/'
```
 

## Managing test subdirectories

Sometimes it is useful to create tests for a package that go beyond simple 
best-practice unit tests. For example you might have some tests that require 
external ressources (web, databases) or take long to execute. If you put such 
tests in subdirectories of `test/testthat` they will neither be automaticaly 
executed by `devtools::test()` nor run on CRAN.

testthis provides helpers do deal with files in subdirectories of 
`tests/testthat`.

* `use_test_subdir()` create a subdir in `tests/testthat/` and 
  (optionally) an R script in `/R` that contains a helper function to run all 
  tests in that subdir. By default, said .R script will be added to your 
  packages `.Rbuildignore` as it is only useful during package development.
  Tests in subdirectories of a packages test dir will *not* be run on CRAN
  on CRAN or by `devtools::test()`.
* You can also just manually use `test_subdir("mysubdir")` to run tests in a
  subdir of `tests/testthat`.
* `test_acceptance()`, `test_manual()` and
  `test_integration()` are presets to run tests in the
  `integration_tests`, `acceptance_tests` and `manual_tests` subdirectories of
  `test/testthat`.


```{r, eval = FALSE}
use_test_subdir("foobar_tests")
## ✔ Creating 'tests/testthat/foobar_tests/'
## creating tester function test_foobar_tests() in r/testthis/R/testthis-testers.R

test_subdir("foobar_tests")
## No tests: no files in r/testthis/tests/testthat/foobar_tests match '^test.*\.[rR]$'
```



## Test coverage checklist

**This is an experimental feature, comments and feature request are welcome**.
Testthis provides a checklist like test coverage analyser. This is 
non-automatic and requires you to manually mark functions as tested in your 
code via special comments. If you look for automatic test coverage analysis, 
you might want to check out the `covr` package instead.

For testthis to recognize wheter a unit test exists for a function, the function
name must either be mention in a `testthat()` `desc` argument, or marked via
special comment tag: `#* @testing functionname`. If you want to mark a function
as not requiring testing, you can add the function name to the file
`/tests/testthat/_testignore` manually, or with `use_testignore()`.


The following example will mark
the functions `parse_testthis_comments` and `detect_testthis_comments` as 
tested:

```{r, eval = FALSE}
# file /tests/testhis/parse_testthis_comments.R

test_that("parse_testthis_comments works as expected", {
  #* @testing detect_testthis_comments
  ...
}
```

The test_coverage summary for a package may look like this (+ marks function for
which tests exist):

```{r}
get_test_coverage()
```




## Testthis Tags

Test_this tags are special comments that modify the behavious of the functions
supplied by this package. They are of the Form `#* @tag <value>`. Please not 
that only some test_this tags really require a `<value>`. The tag system should
be considered to be in a beta-state.

**Valid tags for script files in the /R/ dir (`pkg/R/*.R`)**

* `@testfile <filename>`: manually specifiy associated test file. Should 
  usually start with `test_`. This is used by `test_this()`, `lest_this()` and
  `open_testfile()`.

**Valid tags or test files (`pkg/tests/testthat/test_*.R`)**

* `@testing <functionname>`:  mark `functionname` as tested.
  Should usually go next the associated `test_that()` call. This is used by
  `get_test_coverage()`.
* `@skip`: skip test when running `testthis::test_with_skip()`. Best placed 
  directly below the `context()` call.

also look at `/test/testthat/test_Taglist.R` for examples.
